---
title: "Geography 176A"
author: "[Stephanie Hurtado](https://github.com/shurtadogonzalez/shurtadogonzalez.github.io)"
subtitle: 'Lab 03: COVID-19 Pandemic'
output:
  html_document:
  theme: journal
---

```{r, include=FALSE}

library(tidyverse)


library(knitr)

library(readxl)

library(zoo)

library(kableExtra)
#installing and reading packages 
```

```{r message=FALSE, warning=FALSE}
url = 'https://raw.githubusercontent.com/nytimes/covid-19-data/master/us-counties.csv'
covid = read_csv(url) 
```

# Question 1 

```{r message=TRUE, warning=TRUE}

set_up = covid %>%
  filter(state == "California") %>% 
  group_by(county) %>% 
  mutate(NewCases = cases - lag(cases)) %>% 
  mutate(Cumul_Cases = sum(cases)) %>% 
  ungroup() %>% 
  filter(date == max(date))
```

Table representing the five counties in CA with the most new cases at the most current day
```{r message=TRUE, warning=TRUE}
(most_new_cases = set_up %>% 
  slice_max(NewCases, n = 5) %>% 
  select(county, NewCases))

knitr::kable(most_new_cases,
             caption = "Most New Cases: California Counties",
             col.names = c("County", "New Cases"),
             format.args = list(big.mark = ",")) %>% 
  kableExtra:: kable_styling("basic", full_width =  TRUE, font_size = NULL)
```


Table representing the five counties in CA with the most cumulative cases at the most current day
```{r message=FALSE, warning=FALSE}
(most_cumul_cases = set_up %>% 
    slice_max(Cumul_Cases, n = 5) %>% 
    select(county, Cumul_Cases))

knitr::kable(most_cumul_cases,
             caption = "Most Cumulative Cases: California Counties",
             col.names = c("County", "Cumulative Cases"),
             format.args = list(big.mark = ",")) %>% 
  kableExtra:: kable_styling("basic", full_width =  TRUE, font_size = NULL)
```


```{r, include=FALSE}
library(readxl)

library(formattable)
```

Adding population information 
```{r message=FALSE, warning=FALSE}
PopulationEstimates <- read_excel("../data/PopulationEstimates.xls", skip = 2) %>% 
  dplyr::select(fips = FIPStxt, state = State, county= Area_Name, pop_2019 = POP_ESTIMATE_2019)

```


```{r message=FALSE, warning=FALSE}

pop = PopulationEstimates %>% 
  dplyr::right_join(covid, by = "fips") %>% 
  dplyr::filter(state.y == "California") %>% 
  dplyr::group_by(county.y) %>% 
  mutate(cumul_cases_per_cap = sum(cases) / pop_2019) %>%
  mutate(new_cases_per_cap = ((cases - lag(cases)) /pop_2019)) %>% 
  filter(date == max(date)) %>% 
  ungroup()

```

Table representing the five counties in CA with the most new cases per capita at the most current day
```{r message=FALSE, warning=FALSE}
(most_new_cases_per_cap = pop %>% 
    slice_max(new_cases_per_cap, n = 5) %>% 
    select(county.y, new_cases_per_cap))

knitr::kable(most_new_cases_per_cap,
             caption = "Most New Cases Per Capita: California Counties",
             col.names = c("County", "New Cases Per Capita"),
             format.args = list(big.mark = ",")) %>% 
  kableExtra:: kable_styling("basic", full_width =  TRUE, font_size = NULL)
```

Table representing the five counties in CA with the most cumulative cases per capita at the most current day
```{r message=FALSE, warning=FALSE}
(most_cumul_per_cap = pop %>% 
    slice_max(cumul_cases_per_cap, n = 5) %>% 
    select(county.y, cumul_cases_per_cap))

knitr::kable(most_cumul_per_cap,
             caption = "Most Cumulative Cases Per Capita: California Counties",
             col.names = c("County", "Cumulative Cases Per Capita"),
             format.args = list(big.mark = ",")) %>% 
  kableExtra:: kable_styling("basic", full_width =  TRUE, font_size = NULL)
```

Total number of new cases in the last 14 days per 100,000 people
```{r message=FALSE, warning=FALSE}

last_14_days = covid %>% 
  right_join(PopulationEstimates, by = "fips") %>% 
  filter(date >= max(date, na.rm = TRUE) - 13, state.x == "California") %>% 
  group_by(county.x) %>% 
  summarize(new_cases_per_cap = sum(cases - lag(cases), na.rm = TRUE) / (max(pop_2019) / 100000)) %>% 
  filter(new_cases_per_cap <= 100) %>% 
  ungroup()
```
As of 8/17/2020, there are a total of 14 safe counties in California. This means they comply with the California Department of Public Health’s criteria of having less than 100 new cases per 100,000 residents over the past 14 days.


# Question 2

The number of daily new cases and the 7-day rolling mean of the California, New York, Florida, and Lousiana 
```{r message=FALSE, warning=FALSE}
 
four_states = covid %>% 
  filter (state %in% c("California", "Louisiana", "New York", "Florida")) %>% 
  group_by(state, date) %>%
  summarize(cases = sum(cases)) %>% 
  ungroup() %>% 
  group_by(state) %>% 
  mutate(daily_new_cases= cases-lag(cases), 
         seven_day_roll= zoo::rollmean(daily_new_cases, 7, fill= NA, align= 'right')) %>% 
  ungroup() %>% 
  filter(daily_new_cases > 0)

  ggplot(four_states, aes( x = date)) +
  geom_col(aes(y = daily_new_cases, col= state)) +
  geom_line(aes(y = seven_day_roll, col = "darkred", size = 1)) +
  facet_wrap(~state, scale = "free_y") +
  labs(title = "Daily New Covid Cases at the State Level", x= "", y= "", color = "", caption = "Lab 02", subtitle = "COVID-19 Data: NY Times" ) +
  theme_light() +
  theme(legend.position = 'NA')
```

To get the number of daily new cases per capita and the 7-day rolling mean of the California, New York, Florida, and Lousiana.
```{r message=FALSE, warning=FALSE}
cap_state= PopulationEstimates %>% 
  select(state, pop_2019) %>% 
  right_join(four_states, by = "state") %>% 
  mutate(state_per_cap = ((cases - lag(cases))/ pop_2019), 
        seven_day_roll= zoo::rollmean(daily_new_cases, 7, fill= NA, align= 'right'))
  

ggplot(cap_state, aes(x=date)) +
  geom_col(aes(y= state_per_cap), col= NA, fill= "#F5B8B5") +
  geom_line(aes(y= seven_day_roll), col= "darkred", size= 1) +
  theme_light() +
  labs(title = "States' New Covid Cases Per Capita", x= "", y= "",  caption = "Lab 02", subtitle = "COVID-19 Data: NY Times" ) +
  theme(legend.position = 'NA')

```  

The number of daily, new cases at the state level can be misleading since it doesn't take into consideration the state's population. Once population is integrated into the number of cases, it is expect for data to reflect a more accurate representation of which out of the four states (CA, LA, NY, FL) has been affected more by COVID. 

            
  

  
  

